---
title: "Pilot #2 summary"
output:
  html_document:
    df_print: paged
runtime: shiny
date: "2024-07-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Second pilot summary

The second pilot of the DCE experiment was conducted with the _Intressentpoolen_ panel with 34 individuals participating.
A number of changes were made to the initial survey which was administered among mostly researchers at Lund.
The following changed were made in this pilot test:

* The attribute levels were incorrectly specified in the first pilot test. This was corrected in the second pilot.
* The survey text was shortened to reduce the time to complete the survey.
* Language was refined and reduced to make the survey more easily interpretable.
* Grammatical mistakes resulting from English-Swedish translations were corrected and phrasing was rewritten to make the text more clear.

Respondents were sent a link to the experiment by email.

The experiment is hosted on our private server and connected to our research project domain name:


https://www.reloc-age.org/dce

The survey is powered by the opensource survey framework Formr.




```{r load, echo=FALSE, warning=FALSE, message=FALSE}

library(pacman)

p_load(tidyverse, here, stringr, formr, logitr, cbcTools, texreg, likert, tidyr, haven, kableExtra, sjlabelled, summarytools,gtsummary, sjPlot, sjmisc)



survey_df2 <- readRDS("data/ahrg_pilot_2.rds")
survey_design <- readRDS("data/swe_choice_questions_01.rds")



```

## Descriptive data 

34 individuals from the _Intresspoolen_ participated in the second pilot.
The average monthly housing cost for the participants was 9760 SEK while the average household income was 47000 SEK.
The average planned cost for prospective housing among the the participants was 9820 SEK.
The average time to complete the DCE experiment was 11.3 minutes,
while the average time to complete the feedback section was 3.19 minutes.
Regarding user experience and content, the participants were generally satisfied with an average score of 3.85 out of 5 for user experience and 3.91 out 5 for content.


```{r summary, echo=FALSE, warning=FALSE, message=FALSE}


dfSum <- survey_df2 %>% 
  select(monthcost,
         income,
         planed_cost,
         time_page_1,
         time_page_2,
         time_page_4,
         time_page_feed,
         user_exp = feedback_button01,
         content = feedback_button02,
         mc_position) %>% 
  as.data.frame()

dfSum <- dfSum %>% 
  mutate(
    feed_back_time = time_page_feed/60,
    mc_position1 = factor(mc_position, levels = 1:2,
                         labels = c("Survey first",
                                     "DCE first")
                              )
         ) %>% 
  rowwise() %>%
  mutate(survey_time = sum(time_page_1, time_page_2, time_page_4, na.rm = TRUE)/60)



label(dfSum$monthcost) <- "Monthly costs"
label(dfSum$income) <- "Monthly household income"
label(dfSum$planed_cost) <- "Planned costs"
label(dfSum$user_exp) <- "User Experience"
label(dfSum$content) <- "Content"
label(dfSum$survey_time) <- "Time to complete DCE section"
label(dfSum$feed_back_time) <- "Time to complete feedback"
label(dfSum$mc_position1) <- "Which survey first?"


t1 <- table1::table1(~ monthcost + income + planed_cost + survey_time + feed_back_time + user_exp + content + mc_position1,
                     data = dfSum,
                     na.rm = TRUE,
                     digits = 3,
                     format.number = FALSE,
                     # extra.col=list(`P-value`=pvalue),
                     caption = "Sample description") 

table1::t1flex(t1) |> 
  flextable::fontsize(size = 11) |> 
  flextable::padding(padding = 1, part = "all") 



```

## Regression analysis

While only 34 individuals participated in the second pilot, the data collected from this DCE experiment (34 individuals x 9 choice sets = 306 obs) is sufficient to estimate the preferences of the participants for the attributes of the housing alternatives.

Data results from the survey is analysed using standard methods found in the literature where the utility that an individual assigns to a particular choice alternative is estimated using a conditional logit model based on random utility theory.
The utility of an alternative is represented as:

\begin{equation}
U_{ij} = V_{ij} + \varepsilon_{ij}
\end{equation}

\noindent where $U_{ij}$ is the total utility that respondent \textit{i} associates with choice alternative \textit{j} and $V_{ij}$ is the systematic utility, which is a function of the attributes of the alternative and the preferences of the respondent.
$\varepsilon_{ij}$ is the random error term, representing unobserved factors and individual-specific preferences.
The systematic utility, \( V_{ij} \), is modelled as a linear function of attribute levels:

\begin{equation}
V_{ij} = \beta_1 X_{ij1} + \beta_2 X_{ij2} + \ldots + \beta_k X_{ijk}
\end{equation}

\noindent where $\beta_1$, $\beta_2$, $\ldots$, $\beta_k$  are the coefficients representing the part-worth utilities of the attribute levels and $X_{ij1}$, $X_{ij2}$, $\ldots$, $X_{ijk}$ are the levels of the attributes for choice alternative \textit{j} in scenario \textit{i}.

\vspace{2mm}

\noindent The probability that respondent \( i \) chooses alternative \( j \) from a set of alternatives is modeled using the choice probability as follows:

\begin{equation}
P_{ij} = \frac{e^{V_{ij}}}{\sum_{l=1}^{J} e^{V_{il}}}
\end{equation}

\noindent where $P_{ij}$  is the probability that respondent \textit{i} chooses alternative \textit{j}.
\textit{J} is the total number of alternatives in the choice set.
To estimate the coefficients (\( \beta \) values) in the systematic utility equation, maximum likelihood estimation is used.
The likelihood function for the conditional logit model is maximized to find the best-fitting coefficients that maximize the probability of observing the actual choices made by respondents given the attributes of the alternatives.

Each attribute was coded as a factor variable with the base level set to minimum value for the locational attributes.
The base level for the price attribute was set at zero to estimate the effect of bi directional price changes on the choice of housing.

Below are the results:


```{r regression, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

df_regs2 <- survey_df2 %>% 
  select(respID = respondentID, contains("cbc")) %>% 
  pivot_longer(cols = -respID,
               values_to = "chosen") %>% 
  mutate(qID = extract_numeric(name),
         respID = as.numeric(respID)) %>% 
  filter(!is.na(qID))

df_merged2 <- survey_design %>% 
  inner_join(df_regs2,
             by = c("respID" ,"qID")) %>% 
  left_join(survey_df2 %>% 
              mutate(respondentID = as.numeric(respondentID)
              ),
            by = c("respID" = "respondentID")) %>% 
  mutate(cost = planed_cost * (1+(price/100)),
         choice = if_else(altID == chosen, 1, 0),
         price_con = (1+(price/100)),
         price = factor(price, levels = c("0","-20","-10","10","20")),
         dist_trans = factor(dist_trans, levels = c("800","400","200")),
         dist_green = factor(dist_green, levels = c("15km", "500 meter", "5km")),
         dist_shops = factor(dist_shops, levels = c("15km", "500 meter", "5km")),
  ) %>% 
  select(profileID:parking, choice,monthcost,income, cost,price_con) %>% 
  ungroup() %>% 
  as.data.frame()



m1_dopt2 <- logitr(
  data    = df_merged2 ,
  outcome = "choice",
  obsID   = "obsID",
  pars    = c( "dist_green", "dist_shops","dist_trans","parking", "price")
)



htmlreg(m1_dopt2,
          caption = "Logistic regression results",
          ci.force = TRUE,
          custom.coef.names = c("Distance green - 500 meters (base = 15km)",
                                "Distance green - 5km (base = 15km)",
                                "Distance shops - 500 meters (base = 15km)",
                                "Distance shops - 5km (base = 15km)",
                                "Distance trans - 200m (base = 800m)",
                                "Distance trans - 400m (base = 800m)",
                                "Parking - Garageplats (base = Ingen p-plats)",
                                "Parking - P-plats (base = Ingen p-plats)",
                                "Price -20% (base = 0)",
                                "Price -10% (base = 0)",
                                "Price 10% (base = 0)",
                                "Price 20% (base = 0)"
                                ),
     single.row = TRUE,
     bold = 0.05,
     caption.above = TRUE
        )

```


From the results of the regression analysis, we can see that the distance to green areas, distance to shops, and parking availability all have some significant effects in comparison to their base levels.
All significant coefficients demonstrate the expected signs with positive coefficients representing a increase in utility as distance from the reference level decreases.
Respondents generally prefer to be closer to green areas than shops: they experience a utility change of 1.20 when the distance to green areas is reduced from 15km to 5km, while the utility change is 1.12 when the distance to shops is reduced from 15km to 5km.(__We can compare these as we use the same levels for the attributes__).

Distance to public transport has no significant effect on the choice of housing.
This is probably due to the limited sample size.
__If we did have significant effects, we could not compare them with the other attirbutes as the levels are not standardized.__

Compared to no available parking place, both garage and parking place have a positive effect on the choice of housing.

Regarding price, the results show that a 20% increase in price has a negative effect on the choice of housing, while all other price levels have no significant effect on the choice of housing.
This is probably due to the limited sample size and makes willingness to pay estimates difficult to calculate.

## Comments

Beleow are the comments from the feedback section of the survey. Respondents were asked to provide feedback on the user experience and content of the survey.

### User experience

Delge oss positivia och/eller negativa erfarenheter av att genomföra undersökningen.

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
survey_df2 %>% 
  select(text_feedback2) %>% 
  na.omit %>% 
  sjlabelled::label_to_colnames() %>%
  as_tibble() %>% 
  knitr::kable(caption = "",
               col.names = c("Delge oss positivia och/eller negativa erfarenheter av att genomföra undersökningen")) %>% 
  kable_styling((bootstrap_options = c("striped", "hover"))
  ) %>% 
  row_spec(1:7,
           extra_css = "padding: 12px")
```

### Content


```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
survey_df2 %>% 
  select(text_feedback) %>% 
  na.omit %>% 
  sjlabelled::label_to_colnames() %>%
  as_tibble() %>% 
  knitr::kable(caption = "",
               col.names = c("Delge oss dina tankar om de bostadspreferenser som ingår i undersökningen")) %>% 
  kable_styling((bootstrap_options = c("striped", "hover"))
  ) %>% 
  row_spec(1:7,
           extra_css = "padding: 12px")
```


## Conclusion and proposed changes to the experiment


After receiving the results from the second pilot, a number of considerations have arisen which should be addressed and discussed to develop our experiment further.

* Firstly, while the sample size of 34 individuals is much smaller than our anticipated respondent base of perhaps 600 individuals, we are observing significant effects in the regression analysis. This is a positive sign that the DCE experiment is working as intended and that the preferences of the respondents are being captured by the experiment as we designed it. The number of choice sets per individual (9) and the number of attributes and levels should be more than sufficient to estimate the preferences of the respondents in our target population. __In fact, we should be able to comfortably add another attributes to the experiment to capture more of the preferences of the respondents.__

* Secondly, the regression results demonstrate that further attention to the attribute levels is needed. 
  - There are not many surprises in our results - it is very expected that respondents gain more utility from living closer to amenities.
  - How can we make the results more interpretable and relevant? - __Standardize the levels and create equal intervals for the levels.__
  - Since we used the same levels for distance to shops and green areas, we can compare the coefficient sizes.  We can't compare these attributes to distance to transportation in any meaningful way.
  - Interpreting the coefficients by themselves is difficult since we do not have equal-distant levels for the attributes.  For example, if we consider the distance to shops, we see that respondents experience a utility gain of 1.20 being 14.5km closer to green areas (15km minus 500m) and a utility gain of 1.12 being 10km closer to shops (15km minus 5km).  It is difficult to interpret these figures (14.5km vs. 10km) and a meaningful comparison is challenging as the levels are not equal distant. 
  - __We should standardize the levels of the attributes to make the results more interpretable and to allow for more meaningful comparisons between the attributes.__
  
* Thirdly, the feedback suggests the levels of the attributes may not be applicable to everyone as some people live in rural areas and some urban.  This highlights the need for our presentation of the experiment as a hypothetical scenario.  The scenarios we present do not need to be "one size fits all" and should be simplified to make it easier for the respondents, as well as make our results more interpretable __We should make it clear that the housing choices are hypothetical and that the respondent should consider the attributes as if they were real.__
  - An example of this is to simplify the presentation of the different hypothetical apartments as being located in the city center and changing the attribute distances to be in meters rather than kilometers. Our experiment is by nature hypothetical and does not desire any preconditions to identify preferences. 


## Proposals:

* Make it clear that the housing choices are hypothetical and standardize the choices to "an appartment in the city center" to make the experiment more interpretable and relevant to all respondents.  The wording could be along the lines of : "Imagine you are looking for an apartment in the city center and have to make a choice between the following apartments.  Consider them equal in every way except for the differences contained in the attributes."

* Standardize the levels of the attributes to make the results more interpretable and to allow for more meaningful comparisons between the attributes.  As this experiment would be in the city, our options for all of the attributes could be identical equal distant levels like: 200, 400, and 600 meters. 

* Add another attribute, with the same levels as above, such as "distance to family and friends" - an important attribute identified in our follow up survey.


