---
title: "Discrete choice experiment"
execute: 
  echo: false
  output: asis
  warning: false
  message: false 
format: html
---

## Introduction


```{r} 
#| echo: false
#| message: false
#| warning: false
#| output: false

#Load packages and data
library(pacman)

p_load(tidyverse, here, stringr, formr, logitr, cbcTools, texreg, likert, tidyr, haven, kableExtra,ggrepel, sjlabelled, summarytools,gtsummary)

survey_design <-  read_csv("https://github.com/n-christie/ahrg_dce_survey/blob/main/output/formr/swe_choice_questions_01.csv?raw=true")

survey_df <- readRDS(here("data/formr", "30_09_results.rds")) %>% 
  mutate(ended_survey = if_else(is.na(ended_page_4), ended_page_2, ended_page_4)) %>% 
  filter(created_page_0 > "2024-05-29",
         !is.na(ended_survey))
```



In parallel with the 3-year followup from the RELOC-AGE program, a Discrete Choice Experiment (DCE) was conducted for those respondents choosing to participate.


## Descriptive statistics

The survey was sent out May 26th 2024 with the majority of respondents who completed the DCE doing so directly around this date.
A reminder mail was sent out in early September.
Below shows the number of individuals completing the DCE over the May-October period.

```{r}
survey_df %>%
  ggplot(aes( x = as.Date( created_page_0))) +
  geom_histogram(bins = 150,
                 color = 'black',fill = 'slateblue') +
  scale_x_date(labels = scales::date_format("%b-%d"),
               date_breaks = '2 week') +
  labs(title = "Number of respondents over time",
       x = "",
       y = "Number of respondents")+
  theme_light()


```




Before taking the actual DCE,
respondents were asked to estimate their monthly costs, monthly household income, and planned housing costs.  An opt-out option was included for these financial questions, where a respondent could zero for these questions.
Below presents the descriptive statistics for these variables.





```{r}
#| echo: false

## Descriptive data ----

dfSum <- survey_df %>% 
  # filter(planed_cost != 0,
  #        monthcost != 0,
  #        income != 0) %>%
  select(monthcost,income,planed_cost, time_page_1,time_page_2,time_page_0,time_page_4)%>% 
  mutate(time_page_1 = time_page_1/60,
           cbc_time = (if_else(is.na(time_page_4),time_page_2,time_page_4)/60) ,
         total_survey_time = rowSums(across(c(time_page_1,cbc_time))) ) %>% 
  as.data.frame()


label(dfSum$monthcost) <- "Monthly costs"
label(dfSum$income) <- "Monthly household income"
label(dfSum$cbc_time) <- "Time to complete DCE section (min)"
label(dfSum$total_survey_time) <- "Time to complete entire experiment (min)"
label(dfSum$planed_cost) <- "Planned housing costs"




t1 <- table1::table1(~ monthcost + income + planed_cost + cbc_time + total_survey_time  ,
                     data = dfSum,
                     na.rm = TRUE,
                     digits = 3,
                     format.number = FALSE,
                     # extra.col=list(`P-value`=pvalue),
                     caption = "Sample description") 

table1::t1flex(t1) |> 
  flextable::fontsize(size = 11) |> 
  flextable::padding(padding = 1, part = "all") 
```

It should noted that the "0" values from the respondents who opted not to answer income/cost questions are included in this table.
In this sense, the median values presented are most reflective of the actual distrubutions.

Additionally,
we can see that the time to complete the DCE section, as well as complete the entire experiment is relatively short, with a median of 4.8 minutes to complete the DCE section and 6.3 minutes to complete the entire experiment.

To get a cleaner idea of the values,
I next remove the answers from the respondents who have opted out and generate the following visualizations.


## Visualization of data

```{r}
theme_set(theme_bw())

annotations_p <- data.frame(
  x = c(round(min(dfSum %>% filter(planed_cost != 0) %>% pull(planed_cost)), 2),
        round(mean(dfSum %>% filter(planed_cost != 0) %>% pull(planed_cost)), 2),
        round(max(dfSum %>% filter(planed_cost != 0) %>% pull(planed_cost)), 2)),
  y = c(.000014, .000262, .000015),
  label = c("Min:", "Mean:", "Max:")
) 

planned_plot <- dfSum %>% 
  filter(planed_cost != 0) %>%
  ggplot(aes(x = planed_cost)) +
  geom_histogram(aes(y=..density..), bins = 70,  color = "#000000", fill = "#0099F8")+
  geom_density(color = "#000000", fill = "#F85700", alpha = 0.6) +
  scale_x_continuous(labels = scales::label_number(suffix = " sek")) +
  labs(title = "Planned housing costs",
       y = "",
       x = "") +
    theme(axis.text.y = element_blank()) +
  geom_text(data = annotations_p, aes(x = x, y = y, label = paste(label, x)), size = 3, fontface = "bold")
  
annotations_c <- data.frame(
  x = c(round(min(dfSum %>% filter(monthcost != 0) %>% pull(monthcost)), 2),
        round(mean(dfSum %>% filter(monthcost != 0) %>% pull(monthcost)), 2),
        round(max(dfSum %>% filter(monthcost != 0) %>% pull(monthcost)), 2)),
  y = c(.000014, .000162, .000015),
  label = c("Min:", "Mean:", "Max:")
)

current_plot <-dfSum %>% 
  filter(monthcost != 0) %>%
  ggplot(aes(monthcost)) +
  geom_histogram(aes(y=..density..), bins = 50,  color = "#000000", fill = "#0099F8")+
  geom_density(color = "#000000", fill = "#F85700", alpha = 0.6) +
  scale_x_continuous(labels = scales::label_number(suffix = " sek")) +
  labs(title = "Monthly housing costs",
       y = "",
       x = "") +
  theme(axis.text.y = element_blank()) +
  geom_text(data = annotations_c, aes(x = x, y = y, label = paste(label, x)), size = 3, fontface = "bold")

annotations_i <- data.frame(
  x = c(round(min(dfSum %>% filter(income != 0) %>% pull(income)), 2),
        round(mean(dfSum %>% filter(income != 0) %>% pull(income)), 2),
        round(max(dfSum %>% filter(income != 0) %>% pull(income)), 2)),
  y = c(.000004, .000022262, .000005),
  label = c("Min:", "Mean:", "Max:")
)

income_plot <-dfSum %>% 
  filter(income != 0) %>%
  ggplot(aes(income)) +
  geom_histogram(aes(y=..density..), bins = 50, color = "#000000", fill = "#0099F8")+
  geom_density(color = "#000000", fill = "#F85700", alpha = 0.6) +
  #geom_vline(aes(xintercept = mean(income)),col='red',size=1)+
  scale_x_continuous(labels = scales::label_number(suffix = " sek")) +
  labs(title = "Household income",
       y = "",
       x = "") +
  theme(axis.text.y = element_blank()) +
  geom_text(data = annotations_i, aes(x = x, y = y, label = paste(label, x)), size = 3, fontface = "bold")

annotations_d <- data.frame(
  x = c(round(min(dfSum %>% 
                    filter(planed_cost != 0,
                                   monthcost != 0) %>%
                    mutate(diff_cost = planed_cost - monthcost) %>%
                    pull(diff_cost)), 2),
        round(mean(dfSum %>% 
                   filter(planed_cost != 0,
                          monthcost != 0) %>%
                   mutate(diff_cost = planed_cost - monthcost) %>%
                   pull(diff_cost)), 2),
        round(max(dfSum %>% 
                   filter(planed_cost != 0,
                          monthcost != 0) %>%
                   mutate(diff_cost = planed_cost - monthcost) %>%
                   pull(diff_cost)), 2)),
  y = c(.000014, .000262, .000015),
  label = c("Min:", "Mean:", "Max:")
) 

diff_plot <- dfSum %>% 
  filter(planed_cost != 0,
         monthcost != 0) %>%
  mutate(diff_cost = planed_cost - monthcost) %>% 
  ggplot(aes(x = diff_cost)) +
  geom_histogram(aes(y=..density..), bins = 70,  color = "#000000", fill = "#0099F8")+
  geom_density(color = "#000000", fill = "#F85700", alpha = 0.6) +
  scale_x_continuous(labels = scales::label_number(suffix = " sek")) +
  labs(title = "Difference between planned and current",
       y = "",
       x = "") +
  theme(axis.text.y = element_blank()) +
  geom_text(data = annotations_d, aes(x = x, y = y, label = paste(label, x)), size = 3, fontface = "bold")


gridExtra::grid.arrange(income_plot, current_plot,planned_plot,diff_plot)
```

