---
title: "Discrete Choice Experiments"
format: revealjs
editor: source
---


## For today

We'll take a look at Discrete Choice Experiments and see how they can be used in research.
We'll be touching on such questions as:

::: {.incremental}

- Are Discrete Choice Experiments something I could use?
- Are they hard to implement?
- Why would I even want to use a Discrete Choice Experiment?
- What the heck _is_ a Discrete Choice Experiment?

:::


::: {.notes}
Speaker notes go here.
:::

## The aim today is to:

- Learn about Discrete Choice Experiments - that's a mouthful, let's call them DCEs
- Have a quick background on DCEs - where did they come from?
- See how they work in practice
- Test out a survey
- Analyse the results
- See other ways they are used in practice
- Give you the tools to conduct them yourself.


## Discrete Choice Experiments


## DCE's - the background

```{r}
library(leaflet)
library(tidyverse)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lat = c(55.7047,55.68,55.72), lng = c(13.1910,13.3,13.1),
             popup="The birthplace of R")
  

```


```{r}

```
