---
title: "Discrete Choice Experiments"
format: revealjs
editor: source
---


## For today

We'll take a look at Discrete Choice Experiments and see how they can be used in research.
We'll be touching on such questions as:

::: {.incremental}

- Are Discrete Choice Experiments something I could use?
- Are they hard to implement?
- Why would I even want to use a Discrete Choice Experiment?
- What the heck _is_ a Discrete Choice Experiment?

:::


::: {.notes}
Speaker notes go here.
:::

## The aim today is to:

- Learn about Discrete Choice Experiments - that's a mouthful, let's call them DCEs
- Have a quick background on DCEs - where did they come from?
- See how they work in practice
- Test out a survey
- Analyse the results
- See other ways they are used in practice
- Give you the tools to conduct them yourself.


## Discrete Choice Experiments

- A DCE is a research method used in various fields, including economics, marketing, and healthcare, to understand individuals' preferences and decision-making processes.

- The primary goal of a DCE is to elicit and quantify individuals' preferences for different attributes or features of a product, service, or policy by presenting them with a series of choice scenarios.

::: {.notes}
Key word here is "elicit"
:::

## Choice experiments

- DCE's fall into the more broader group of choice experiments.
- When we make a choice, we show a preference.
- Preferences can be construed as __stated__ or __revealed__.

## Stated preferences

- Definition: Stated preferences refer to the preferences that individuals explicitly state or declare in response to direct questioning or surveys.

- Methodology: Researchers gather information by directly asking individuals about their preferences, opinions, or choices through interviews, surveys, or other data collection methods.

- Applications: Stated preferences are commonly used in market research, policy analysis, and product development to understand consumer choices and inform decision-making processes.


## Revealed preferences

- Definition: Revealed preferences are inferred from individuals' actual behaviors and choices, providing insights into their true preferences based on observed actions rather than stated declarations.

- Methodology: Researchers analyze real-world data, such as consumer purchasing behavior, travel patterns, or observed choices, to uncover implicit preferences.

- Advantages: Revealed preferences offer a more accurate reflection of individuals' preferences as they are derived from observable actions, reducing biases associated with self-reporting.

## Discrete Choice Experiments

- DCE's are a type of revealed preference experiments. 
- By asking respondents to choose between alternatives in a controlled experiment we can elicit preferences from the sample.

## HTML WIDGETS

```{r}
library(leaflet)
library(tidyverse)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lat = c(55.7047,55.68,55.72), lng = c(13.1910,13.3,13.1),
             popup="The birthplace of R")
  

```

## Plotly
```{r, echo=FALSE, results='asis', fig.height=4}
library(ggplot2)
library(plotly)
p <- ggplot(data = diamonds, aes(x = cut, fill = clarity)) +
            geom_bar(position = "dodge")
ggplotly(p)
```


## Diagramme R
```{r, echo=FALSE, results='asis', fig.height=4}
library(DiagrammeR)
grViz("
  digraph {
    layout = twopi
    node [shape = circle]
    A -> {B C D} 
  }")
```


## Dygraphs
```{r, echo=FALSE, results='asis'}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```
