---
title: "Discrete Choice Experiments"
format: 
  revealjs:
    multiplex: true
    self_contained: true
editor: source
---

# Discrete choice experiments

We'll take a look at Discrete Choice Experiments and how we are implementing one in the upcoming follow-up survey. We'll be touching on such questions as:

::: incremental
-   Are Discrete Choice Experiments something I could use?
-   How do you conduct a DCE?
-   Why would I even want to use a Discrete Choice Experiment?
-   What the heck ***is*** a Discrete Choice Experiment?
-   How are we using them in our survey?
-   How we can make our survey better
:::

::: notes
We'll cover some best practices and need more feedback!
:::

## Today's aim:

::: incremental
-   Learn about Discrete Choice Experiments - that's a mouthful, let's call them DCEs
-   Have a quick background on DCEs - where did they come from?
-   Attributes and levels - things to consider
-   Participate in an updated survey
-   Feedback and discussion
:::

# Discrete Choice Experiments

-   A DCE is a research method used in various fields, including economics, marketing, and healthcare, to understand individuals' preferences and decision-making processes.

-   The primary goal of a DCE is to elicit and quantify individuals' preferences for different attributes or features of a product, service, or policy by presenting them with a series of choice scenarios.

::: notes
Key word here is "elicit"
:::

## Choice experiments

-   DCE's fall into the more broader group of choice experiments.
-   When we make a choice, we show a **preference**.
-   Preferences can be construed as **stated** or **revealed**.

## Stated vs. revealed preferences

-   Stated: Researchers gather information by directly asking individuals about their preferences, opinions, or choices through interviews, surveys, or other data collection methods.

-   Revealed: Researchers analyze real-world data, such as consumer purchasing behavior, travel patterns, or **observed choices**, to uncover implicit preferences.

## Revealed preferences - advantages

-   Advantages: Revealed preferences offer a more accurate reflection of individuals' preferences as they are derived from observable actions, reducing biases associated with self-reporting.

## Discrete Choice Experiments

-   DCE's are a type of revealed preference experiments.
-   By asking (forcing) respondents to choose between alternatives in a controlled experiment we can elicit preferences from the sample.

## The experiment

The



# Slide


::: incremental
-
-
-
-

:::

::: notes
Speaker notes go here
:::



# RELOC-AGE

-   Around 1900 Respondents across Sweden

```{r map, eval=TRUE ,warning=FALSE, include=TRUE}
library(leaflet)

address_lat_lon <- readRDS("discrete_choice_experiments_files/coords.rds") 

leaflet(address_lat_lon,width = "100%" ) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(~lon, ~lat, popup="The birthplace of R",
             clusterOptions = markerClusterOptions())
```

## Aim and RQ's

-   "The overarching aim of this paper is to explore heterogeneous housing preferences of those aged 55 years or more and interested in relocation"
-   

## Attributes and levels

## HTML WIDGETS

```{r}
library(leaflet)
library(tidyverse)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lat = c(55.7047,55.68,55.72), lng = c(13.1910,13.3,13.1),
             popup="The birthplace of R")
  

```

## Plotly

```{r, echo=FALSE, results='asis', fig.height=4}
library(ggplot2)
library(plotly)
p <- ggplot(data = diamonds, aes(x = cut, fill = clarity)) +
            geom_bar(position = "dodge")
ggplotly(p)
```

## Diagramme R

```{r, echo=FALSE, results='asis', fig.height=4}
library(DiagrammeR)
grViz("
  digraph {
    layout = twopi
    node [shape = circle]
    A -> {Abb C D} 
  }")
```

## Dygraphs

```{r, echo=FALSE, results='asis'}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```
